---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DemografixeR

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/matbmeijer/DemografixeR.svg?branch=master)](https://travis-ci.org/matbmeijer/DemografixeR)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/matbmeijer/DemografixeR?branch=master&svg=true)](https://ci.appveyor.com/project/matbmeijer/DemografixeR)
<!-- badges: end -->

DemografixeR is an API wrapper of all 3 Demografix API's - all three APIs supported in one package:

* [https://genderize.io/](https://genderize.io/) - **Gender estimation** based on a name
* [https://agify.io/](https://agify.io/) - **Age estimation** based on a name
* [https://nationalize.io/](https://nationalize.io/) - **Nationality estimation** based on a name

## Documentation

You can find all the necessary documentation about the package here:
[https://matbmeijer.github.io/DemografixeR/](https://matbmeijer.github.io/DemografixeR/)

## Installation

You can install the released version of DemografixeR following the next commands:

``` r
if (!require("devtools")) install.packages("devtools")
devtools::install_github("matbmeijer/DemografixeR")
```

## Examples

These are basic examples which shows you how to estimate **nationality**, **gender** and **age** by a given name with & without specifying a country. It also shows that the functions are able to deal with missing values:

```{r example}
library(DemografixeR)

#Simple example without country_id
names<-c("Ben", "Allister", "Lucie", "Paula")
genderize(name = names)
nationalize(name = names)
agify(name = names)

#Simple example with
genderize(name = names, country_id = "US")
agify(name = names, country_id = "US")

#Workflow example with dplyr with missing values and multiple different countries
df<-data.frame(names=c("Ana", NA, "Pedro",
                       "Francisco", "Maria", "Elena"),
                 country=c(NA, NA, "ES",
                           "DE", "ES", "NL"), stringsAsFactors = FALSE)

df %>% dplyr::mutate(guessed_nationality=nationalize(name = names),
                guessed_gender=genderize(name = names, country_id = country),
                guessed_age=agify(name = names, country_id = country)) %>% 
  knitr::kable()

#Detailed data.frame example:
genderize(name = names, simplify = FALSE, meta = TRUE) %>% knitr::kable()

```

Please note that the 'DemografixeR' project is released with a
  [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By contributing to this project, you agree to abide by its terms.
